{\bf Practical Concurrency Testing} \\ \normalsize \vspace{1em} \begin{tabular}{c} {\em or: How I Learned to Stop Worrying and Love the Exponential Explosion} \end{tabular}

### 2018-12-06 thesis/landslide.tex
adjust placement of sleepsets figure
landslide summary

\subsubsection{Design}

@@

\begin{itemize}
	\item {\tt mem\_check\_shared\_access()} ({\tt memory.c})
		will assert that kernel memory is direct-mapped.
	\item {\tt use\_after\_free()} ({\tt memory.c})
		will ignore use-after-free reads originating from kernel code during the {\tt swexn} system call
		(an extremely common and neither harmful nor technically interesting bug among student implementations).
	\item {\tt cause\_test()} ({\tt test.c})
		will issue keyboard input to type the test case name and press enter
		when the initialization sequence completes and the shell is blocked on {\tt readline}.
	\item {\tt kern\_mutex\_block\_others()} ({\tt schedule.c})
		will handle the special ``blocked on via mutex'' state changes
		whenever a mutex is acquired or released,
		for kernels which use the {\tt tell\_landslide\_kern\_mutex\_blocking()} annotation.
	\item {\tt sched\_update\_kern\_state\_machine()} ({\tt schedule.c})
		will handle the reference kernel's invocation of {\tt sched\_unblock()} within {\tt cond\_\allowbreak{}signal()}
		as a signal event for happens-before analysis.
	\item {\tt cause\_timer\_interrupt\_immediately()} ({\tt x86.c})
		will read the {\tt esp0} value out of the TSS to support user-to-kernel mode switches
		in timer interrupts injected during userspace execution.
	\item {\tt splice\_pre\_vanish\_trace} ({\tt stack.c})
		will, when a {\tt vanish}ing thread has already deleted its Simics symbol table object,
		splice in a saved ``pre-vanish'' stack trace
		(saved previously in {\tt sched\_update\_kern\_state\_machine()})
		so the user can see the userspace execution sequence preceding the {\tt vanish} invocation.
	\item {\tt stack\_trace} ({\tt stack.c})
		will, when {\tt ebp} crosses from kernel- to userspace across a system call boundary
		(a reference kernel feature to allow Simics stack traces to cross same),
		use {\tt PATHOS\_SYSCALL\_IRET\_DISTANCE} (\cref{sec:landslide-config-landslide})
		to track {\tt esp}'s value across the stack switch.
\end{itemize}

@@

@@

\begin{figure}[h]
	\begin{center}
		\begin{tabular}{p{0.30\textwidth}p{0.30\textwidth}p{0.30\textwidth}}
	\begin{center}
	\begin{tabular}{l}
		\dporTAcode{tmp1 <- x (read)} \\
		\dporTBcode{tmp2 <- x (read)} \\
		\dporTAcode{x <- tmp1 (write)} \\
		\dporTBcode{x <- tmp2 (write)} \\
		% old example, pre dpor-example figure
		%$T_1$@$t_a$: \texttt{write~x} \\
		%$T_2$@$t_b$: \texttt{read~y} \\
		%$T_2$@$t_c$: \texttt{read~x} \\
	\end{tabular}
	\end{center}
	&
	\begin{center}
	\begin{tabular}{l}
		\dporTBcode{tmp2 <- x (read)} \\
		\dporTBcode{x <- tmp2 (write)} \\
		\dporTAcode{tmp1 <- x (read)} \\
		\dporTAcode{x <- tmp1 (write)} \\
		%$T_2$@$t_{b'}$: \texttt{read~y} \\
		%$T_2$@$t_{c'}$: \texttt{read~x} \\
		%$T_1$@$t_{a'}$: \texttt{write~x} \\
	\end{tabular}
	\end{center}
	&
	\begin{center}
	\begin{tabular}{l}
		\dporTBcode{tmp2 <- x (read)} \\
		\dporTAcode{tmp1 <- x (read)} \\
		\dporTAcode{x <- tmp1 (write)} \\
		\dporTBcode{x <- tmp2 (write)} \\
		%$T_2$@$t_{b'}$: \texttt{read~y} \\
		%$T_1$@$t_{a''}$: \texttt{write~x} \\
		%$T_2$@$t_{c''}$: \texttt{read~x} \\
	\end{tabular}
	\end{center}
	\\
	\begin{center}
	(a) Original branch (C).
	\end{center}
	&
	\begin{center}
	(b) Goal branch (K).
	\end{center}
	&
	\begin{center}
	(c) Redundant branch (L).
	\end{center}
	\end{tabular}
	\end{center}
	\caption[Motivating example for the sleep sets optimization.]
		{Motivating example for the sleep sets optimization.
	Three of \Cref{fig:dpor-example-2}'s interleavings are highlighted,
	with the always-independent {\tt tmpN++}s omitted for brevity.
	}
	\label{fig:sleepsets}
\end{figure}

@@

>>>>

\subsubsection{Overview}

@@

\begin{itemize}
	\item {\tt mem\_check\_shared\_access()} ({\tt memory.c})
		will assert that kernel memory is direct-mapped.
	\item {\tt use\_after\_free()} ({\tt memory.c})
		will ignore use-after-free reads originating from kernel code during the {\tt swexn} system call
		(an extremely common and neither harmful nor technically interesting bug among student implementations).
	\item {\tt cause\_test()} ({\tt test.c})
		will issue keyboard input to type the test case name and press enter
		when the initialization sequence completes and the shell is blocked on {\tt readline}.
	\item {\tt kern\_mutex\_block\_others()} ({\tt schedule.c})
		will handle the special ``blocked on via mutex'' state changes
		whenever a mutex is acquired or released,
		for kernels which use the {\tt tell\_landslide\_kern\_mutex\_blocking()} annotation.
	\item {\tt sched\_update\_kern\_state\_machine()} ({\tt schedule.c})
		will handle the reference kernel's invocation of {\tt sched\_unblock()} within {\tt cond\_\allowbreak{}signal()}
		as a signal event for happens-before analysis.
	\item {\tt cause\_timer\_interrupt\_immediately()} ({\tt x86.c})
		will read the {\tt esp0} value out of the TSS to support user-to-kernel mode switches
		in timer interrupts injected during userspace execution.
	\item {\tt splice\_pre\_vanish\_trace} ({\tt stack.c})
		will, when a {\tt vanish}ing thread has already
		\revisionminor{unmapped all of its user memory,}
		%deleted its Simics symbol table object,
		splice in a saved ``pre-vanish'' stack trace
		(saved previously in {\tt sched\_update\_kern\_state\_machine()})
		so the user can see the userspace execution sequence preceding the {\tt vanish} invocation.
	\item {\tt stack\_trace} ({\tt stack.c})
		will, when {\tt ebp} crosses from kernel- to userspace across a system call boundary
		(a reference kernel feature to allow Simics stack traces to cross same),
		use {\tt PATHOS\_SYSCALL\_IRET\_DISTANCE} (\cref{sec:landslide-config-landslide})
		to track {\tt esp}'s value across the stack switch.
\end{itemize}

@@

% TODO: check figure placement
% should be right below dpor-example-2, on the previous page
% i had to put ! not just ht to get it to show up correctly
\begin{figure}[ht!]
	\begin{center}
		\begin{tabular}{p{0.30\textwidth}p{0.30\textwidth}p{0.30\textwidth}}
	\begin{center}
	\begin{tabular}{l}
		\dporTAcode{tmp1 <- x (read)} \\
		\dporTBcode{tmp2 <- x (read)} \\
		\dporTAcode{x <- tmp1 (write)} \\
		\dporTBcode{x <- tmp2 (write)} \\
		% old example, pre dpor-example figure
		%$T_1$@$t_a$: \texttt{write~x} \\
		%$T_2$@$t_b$: \texttt{read~y} \\
		%$T_2$@$t_c$: \texttt{read~x} \\
	\end{tabular}
	\end{center}
	&
	\begin{center}
	\begin{tabular}{l}
		\dporTBcode{tmp2 <- x (read)} \\
		\dporTBcode{x <- tmp2 (write)} \\
		\dporTAcode{tmp1 <- x (read)} \\
		\dporTAcode{x <- tmp1 (write)} \\
		%$T_2$@$t_{b'}$: \texttt{read~y} \\
		%$T_2$@$t_{c'}$: \texttt{read~x} \\
		%$T_1$@$t_{a'}$: \texttt{write~x} \\
	\end{tabular}
	\end{center}
	&
	\begin{center}
	\begin{tabular}{l}
		\dporTBcode{tmp2 <- x (read)} \\
		\dporTAcode{tmp1 <- x (read)} \\
		\dporTAcode{x <- tmp1 (write)} \\
		\dporTBcode{x <- tmp2 (write)} \\
		%$T_2$@$t_{b'}$: \texttt{read~y} \\
		%$T_1$@$t_{a''}$: \texttt{write~x} \\
		%$T_2$@$t_{c''}$: \texttt{read~x} \\
	\end{tabular}
	\end{center}
	\\
	\begin{center}
	(a) Original branch (C).
	\end{center}
	&
	\begin{center}
	(b) Goal branch (K).
	\end{center}
	&
	\begin{center}
	(c) Redundant branch (L).
	\end{center}
	\end{tabular}
	\end{center}
	\caption[Motivating example for the sleep sets optimization.]
		{Motivating example for the sleep sets optimization.
	Three of \Cref{fig:dpor-example-2}'s interleavings are highlighted,
	with the always-independent {\tt tmpN++}s omitted for brevity.
	}
	\label{fig:sleepsets}
\end{figure}

@@

@@

\section{Summary}

\revision{This chapter has presented Landslide,
a stateless model checker for Pebbles and Pintos kernels and Pebbles userspace thread libraries.
For the most part, Landslide's contribution to the world of concurrency testing is as a feat of engineering
which combines many existing techniques from prior work in a single implementation.
These include:

\begin{itemize}
	\item Uses Dynamic Partial Order Reduction (DPOR) \cite{dpor}
		to prune equivalent interleavings while exploring the state space (\cref{sec:landslide-dpor}).
	\item Uses its simulator's memory tracing integration \cite{simics,bochs}
		to power data race analysis \cite{djit,fasttrack} (\cref{sec:landslide-datarace}).
		%and DPOR \cite{dpor} (\cref{sec:landslide-dpor}).
	\item Supports Preemption Sealing \cite{sealing} to configure the set of active preemption points
		and hence keep state space sizes manageable (\cref{sec:landslide-pps}).
	\item Extends DPOR with Recursive and Weighted Backtrack Estimation \cite{estimation}
		to predict overall ultimate state space size and completion time (\cref{sec:landslide-estimate}).
	\item Extends DPOR (at the user's option) with Iterative Context Bounding \cite{chess-icb}
		to heuristically uncover bugs faster should they exist,
		at the expense of overall completion time (\cref{sec:landslide-icb}).
\end{itemize}

However, over the years Landslide has also grown some convenience and/or accessibility features
that could be considered contributions in their own right.

\begin{itemize}
	\item Extends Preemption Sealing for finer-grained control
		over whether individual functions and/or global variables
		should be considered for, or excluded from,
		DPOR and/or data-race analysis
		(\cref{sec:landslide-config-landslide};
		{\tt thrlib\_function} and {\tt ignore\_dr\_function}, respectively).
		This allows a test case author to precisely target the exact subset of code
		she is interested in testing,
		avoiding unnecessary state space inflation from trusted code.
	\item Adds a mutex testing mode to configure whether the
		data race analysis should consider lock' internal accesses
		(\cref{sec:landslide-dynamicconfig}; {\tt TESTING\_MUTEXES}).
		This allows the user to verify that her lock implementation
		correctly provides mutual exclusion
		(upcoming in \cref{sec:education-pebbles-tests}; {\tt mutex\_test}),
		before relying on them as trusted code in subsequent tests.
	\item Implements the sleep set extension to DPOR \cite{partial-order-methods,dpor,optimal-dpor},
		which identifies and skips further equivalences based on which memory accesses
		DPOR intended to reorder,
		%, rather than just switching threads and seeing what happens
		in a way that is slightly less optimal compared to prior work,
		but much simpler to implement
		and requires less state at runtime (\cref{sec:landslide-sleepsets}).
	\item Heuristically identifies infinite loops and livelocks
		by leveraging knowledge of how long the test program should be expected to run
		obtained during previous executions,
		uniquely suited to stateless model checking (\cref{sec:landslide-infloop}).
	\item Provides automatic code instrumentation facilities,
		beyond just the simple task of identifying important function boundaries,
		that attempt to understand a wide variety of even the most unusual student implementations
		(\cref{sec:landslide-pebblesspecifics},
		\cref{sec:landslide-pintosspecifics}).
		This includes an advanced heuristic for detecting ad-hoc synchronization loops
		that Landslide should treat as blocking, %a thread,
		even if it does not trigger the automatically-inserted blocking annotations
		(\cref{sec:landslide-blocking-yield}),
		and a further heuristic to ensure that any inaccuracy on the part of the first heuristic
		does not result in false-positive bug reports (\cref{sec:landslide-fp-deadlock}).
\end{itemize}

The following three chapters, \Cref{chap:quicksand,chap:education,chap:tm},
will each build upon Landslide to present the thesis's main research contributions.}
